---
title: "STA130 Rstudio Homework"
author: "[Student Name] ([Student Number]), with Josh Speagle & Scott Schwartz"
subtitle: OPTIONAL Problem Set 4.5 (UNGRADED)
urlcolor: blue
output:
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE, message=TRUE, warning=FALSE)
```

# Instructions

Complete the exercises in this `.Rmd` file and submit your `.Rmd` and knitted `.pdf` output through [Quercus](https://q.utoronto.ca/courses/296457) by 11:59 pm E.T. on Thursday, February 9.

**Please note that this supplementary set of R exercises is optional but recommended. It will not be graded.**

```{r, message=FALSE}
library(tidyverse)
```

```{r}
student_num =   # list your student number
student_num_last3 =   # list the last three digits of your student number
```


## Question 1: Specifying Null Hypotheses

In class so far, w have covered two kinds of hypothesis tests. 

- In the first version, we hypothesize a fraction in a population (the parameter $f_{\rm pop}$) and test that value using the fraction observed in the sample (what we have called $f_{\rm samp}$). These *one sample hypotheses tests* have a null hypothesis of the form $H_0: f_{\rm samp} = f_{\rm pop}$, with an alternative hypothesis being $H_1: f_{\rm samp} \neq f_{\rm pop}$ (2-sided) or $H_1: f_{\rm samp} > f_{\rm pop}$ or $H_1: f_{\rm samp} < f_{\rm pop}$ (1-sided). Since this one-sample framework works for any test statistic, a more general form of this test that is often encountered is $H_0: T_{\rm samp} = T_{\rm pop}$, often with $T$ being something like the "center" (mean or median).

- In the second version, we hypothesize a relationship between two populations, such as that both populations have the same test statistic (mean, median, proportion, standard deviation, interquartile range, etc.). These *two sample hypotheses tests* have a null hypothesis of the form $H_0: T_1 = T_2$ (alternately, $H_0: T_1 - T_2 = 0$), with the alternative hypothesis defined the same way as above (i.e. $H_1: T_1 > T_2$, $H_1: T_1 < T_2$, or $H_1: T_1 \neq T_2$).

For each of the following scenarios, state appropriate null and alternative hypotheses $H_0$ and $H_1$. Be sure to carefully define any parameters you refer to. For each scenario, also state in simple terms what the claim of the null hypothesis is. 

\
**(a)** A health survey asked individuals to report the number of times they exercised each week. Researchers were interested in determining if the proportion of individuals who exercised at least 100 minutes per week differed between people who live in the condos vs people who do not live in condos.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(b)** A study was conducted to examine whether a baby is born prematurely/early (i.e. before their due date) to whether or not the baby's mother smoked while she was pregnant.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(c)** Nintendo is interested in whether or not their online advertisements are working. They record whether or not a user had seen an ad on a given day and their amount of spending on Nintendo products in the next 48 hours. They are interested in determining if there is an association between whether or not the user saw an ad and their short-term future expenditures.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(d)** Based on results from a survey of graduates from the University of Toronto, we would like to compare the median salaries of graduates from statistics programs versus graduates from mathematics programs.

> *REPLACE THIS TEXT WITH YOUR ANSWER*


# Question 2: Fisher's Tea Experiment

There is an interesting [account](https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2012.00620.x) of the British statistician Ronald Fisher at a tea party in the 1920s. One of the other guests was algae scientist Dr. Muriel Bristol, who refused a cup of tea from Fisher because he put milk in first BEFORE pouring the tea. Bristol was convinced she could taste the difference, and much preferred the taste of tea where the milk was poured in afterwards. Fisher didn't think that there could be a difference and proposed a hypothesis test to examine the situation.

Fisher made 8 cups of tea, 4 with milk in first and 4 with tea in first, and gave them to Dr. Bristol without her seeing how they were made and she would say if she thought the tea or the milk was poured first.  As it turned out, Dr. Bristol correctly identified if the tea or milk was poured first for all 8 of the cups. Fisher, being a skeptical statistician wanted to test if this could be happening by chance with Bristol just randomly guessing (or whether it seemed more likely that Bristol was not guessing). 

Suppose you run an experiment like this with students in STA130. You prepare 40 British-style cup of teas with milk in first and 40 with tea in first. You then get a random sample of 80 STA130 students to each taste one cup and tell you whether they think the milk or tea was poured first. 49 students correctly state which was poured first.

**(a)** Write down a hypothesis test to determine whether the students are just randomly guessing or not (i.e whether they are performing better or worse than expected).

*Hint: Think carefully about whether you are performing a 1-sided or 2-sided hypothesis test here.*

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(b)** Carry out a simulation study and plot the simulated sampling distribution under the null hypothesis $H_0$. Then, compute the $p$-value given that 49 students correctly stated whether the milk or tea was poured first.

*Hint: You should be able to re-use a lot of your code from Question 3 in HW4 here!*

```{r}
set.seed(student_num_last3 + 2)  # REQUIRED so the results are reproducible!
N <- 1000  # feel free to increase this is you want finer precision

# Code your answers here

```

\
**(c)** Given an $p < \alpha=0.05$ rejection rule, what is your conclusion about this hypothesis test based on your computed $p$-value?

> *REPLACE THIS TEXT WITH YOUR ANSWER*


# Question 3: Airbags

The table below is adapted from *"Biostatistics for the Biological and Health Sciences"* and presents data from a random sample of passengers sitting in the front seat of cars involved in car crashes. Based on this data we'd like to make a determination as to whether or not death rates differ for passengers in cars with airbags and passengers in cars without airbags.

|                           | Airbag available | No airbag available |
|---------------------------|------------------|---------------------|                           
| Passenger Fatalities      |  45              | 62                  |
| Total number of Passengers|  10,541          | 9,867               |

The code below creates a tidy data frame for this problem using the `rep()` function.

```{r}
data <- tibble(group = c(rep("airbag", 10541), rep("no_airbag", 9867)),
               outcome = c(rep("dead", 45), rep("alive", 10541 - 45), 
                           rep("dead", 62), rep("alive", 9867 - 62)))
```

\
**(a)** What is the null hypotheses $H_0$ in terms of $f_{\rm airbag}$ and $f_{\rm nobag}$? What is the alternative hypothesis $H_1$ in terms of $H_0$? In a simple sentence without using $H_0$, $f_{\rm airbag}$, and $f_{\rm nobag}$ notation, what is the claim of the null hypothesis?

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(b)** Compute the test statistic. Then, carry out a simulation study under a permutation test and plot the simulated sampling distribution under the null hypothesis $H_0$. Finally, compute the $p$-value given the value of the test statistic.

*Hint: You should be able to re-use a lot of your code from Question 4 here!*

```{r}
# compute the test statistic below

```

```{r}
set.seed(student_num_last3 + 3)  # REQUIRED so the results are reproducible!

# Code your answers here  for the simulation study

```

\
**(c)** At the **$\alpha=0.01$ significance level** (not the $\alpha=0.05$ significance level!), what is your conclusion about this hypothesis test based on the $p$-value computed above?

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(d)** Based on your conclusion above, what kind of error could you have made and why?

(A) Type I error

(B) Type II error

(C) Both

> *REPLACE THIS TEXT WITH YOUR ANSWER*

\
**(e)** Does your conclusion support the claim that "airbags save lives"? Or does it seem reasonable to believe that there could be some sort of confounding variable (like in Question 4 of HW4)? For instance, that people who choose to drive in cars without airbags have different driving habits than those who choose to drive in cars with airbags? Or perhaps people who drive in cars without airbags drive in very different environments than people who drive in cars with airbags?

> *REPLACE THIS TEXT WITH YOUR ANSWER*