---
title: "STA130 Rstudio Homework"
author: "[Student Name] ([Student Number]), with Josh Speagle & Scott Schwartz"
subtitle: Problem Set 2
urlcolor: blue
output:
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE, message=TRUE, warning=FALSE)
```

# Instructions

Complete the exercises in this `.Rmd` file and submit your `.Rmd` and knitted `.pdf` output through [Quercus](https://q.utoronto.ca/courses/296457) by 11:59 pm E.T. on Thursday, January 26. 

# Question 1: Data Types

Problem Set 1 included the following code:

```{r, eval=FALSE, echo=TRUE}
my_answers <- c(r1,r2,c1,c2)
square_answers <- c(10,-1,3,12)
```

For the first three questions below, choose the correct answer from the following statements:

(1) A single value counting how many correct rows and columns you calculated.
(2) A numeric vector of the differences between the math square answers and your answers (should be all 0s if you got them all right).
(3) A character vector of 'TRUE' and 'FALSE', 'TRUE' for each answer that matches and 'FALSE' for any that don't.
(4) A logical vector of `TRUE` and `FALSE`, `TRUE` for each answer that matches and `FALSE` for any that don't.
(5) A single logical value `TRUE` or `FALSE`, `TRUE` if all the values match, `FALSE` if any of the values don't match.

**(a)** Which of the above best describes what `my_answers == square_answers` is?

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(b)** Which of the above best describes what `sum(my_answers == square_answers)` is?

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(c)** Which of the above best describes what `all(my_answers == square_answers)` is?

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(d)** What is the sequence of steps involved in getting the answer for `sum(c(TRUE,FALSE))`? What additional step is required to get the answer for `sum(my_answers == square_answers)`?

> *REPLACE THIS TEXT WITH YOUR ANSWER*

*Hint: The `sum` function works only on `numeric` data types and does not itself directly know anything about `logical` data types. How might this relate to the concept of **coercion**?*

# Question 2: Super Bowl Ads

The data for this question will be based on a sample of Super Bowl ads. This is stored in the file `superbowl_ads.csv` in the same directory as this file and includes the following variables:

- `year` (double) Superbowl year
- `brand`	(character)	Brand for commercial
- `funny`	(logical)	Contains humor
- `show_product_quickly`	(logical))	Shows product quickly
- `celebrity`	(logical) Contains celebrity
- `danger`	(logical)	Contains danger
- `view_count`	(double)	Youtube view count
- `like_count`	(double)	Youtube like count
- `dislike_count`	(double)	Youtube dislike count
- `superbowl_ads_dot_com_url`	(character)	Superbowl ad URL
 
*This data was posted on [GitHub](https://github.com/fivethirtyeight/superbowl-ads#super-bowl-ads) by the data-oriented reporting outlet [FiveThirtyEight](https://github.com/fivethirtyeight) and subsequently featured on [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-02/readme.md). For more information, see the above links.*

```{r, message=FALSE}
library(tidyverse) # Load the tidyverse functionality so it is available to use 
superbowl <- read_csv("superbowl_ads.csv")
```

**(a)** Use the `glimpse()` function to view the properties of the `superbowl` data set. How many rows and columns are there? How many observations does it include? How many variables are measured for each observation? 

```{r}
# replace this with your code
```

*REPLACE THIS TEXT WITH YOUR ANSWER*

**(b)** Explore the distribution of `view_count` using a histogram with 2 bins, a histogram with 8 bins, and a histogram with 50 bins (3 histograms total). Make sure to specify meaningful axis labels where appropriate. 

```{r}
# You can put multiple plots in the same code chunk

```

```{r}
# Or your can put different plots in separate code chunks

```

```{r}
# Feel free to add or remove code chunks as desired

```

Which of these histograms is most appropriate to describe the distribution of `view_count`? Why? Write a few sentences describing the distribution based on the histogram you chose as most appropriate.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(c)** Construct two plots (your choice) to visualize the distribution of `brand` and one of the following other categorical variables from the superbowl ads data: `funny`, `danger`, or `celebrity`. Make sure to specify meaningful axis labels where appropriate.

*Hint: If you choose a categorical variable with many different categories, you may find it useful to use `coord_flip()` to flip the bars horizontally and/or change the options in the R code chunk to make the plot larger (e.g., {r, fig.height=15, fig.width=5}).*

```{r, message=FALSE, fig.height=3, fig.width=4, fig.align='center'}
# One reason to use use different code chunks for different figures is
# to assign different figure aspect ratio controls to different figure

```

Describe the distribution in 1-2 sentences.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(d)** Construct a joint set of two boxplots showing visual summaries of the distribution of number of likes (`like_count`) depending on whether ads included a celebrity or not (`celebrity`). Make sure to specify meaningful axis labels where appropriate.

```{r}
# This should be a single plot, NOT TWO! 
# Boxplots can be put in the same plot!

```

Write 2-4 sentences comparing these distributions.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

# Question 3: Births and Smoking

The `births` data set is part of the `openintro` package. It consists of random sample of 100 births for babies in North Carolina where the mother was not a smoker and another 50 where the mother was a smoker. The code below loads the required libraries for this question and provides a glimpse of the `births` data frame.

*Hint: Type `?births` in the R console for more information about the data.*

```{r, echo=FALSE}
library(tidyverse)
library(openintro)
```

```{r}
births %>% glimpse()
```
  
**(a)** Choose two categorical variables from the `births` data set and plot the distribution of each one in separate plots using the visualization method of your choice.

```{r, message=FALSE, fig.height=3, fig.width=4, fig.align='center'}
# replace this with your plot
```

Identify whether each of these variables is a nominal or ordinal categorical variable and write one or two sentences interpreting each plot.  

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(b)** Choose a quantitative variable from the `births` data set and plot its distribution using the visualization method of your choice.

```{r, message=FALSE, fig.width=3, fig.height=4, fig.align='center'}
# replace this with your plot
```

Identify whether the variable you selected is continuous or discrete, and write 2-3 sentences describing the distribution.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(c)** Construct a plot that shows the relationship between birth weight (`weight`) and mother's smoking status (`smoke`). Make sure to specify meaningful axis labels where appropriate.

```{r, message=FALSE, fig.width=3, fig.height=4, fig.align='center'}
# replace this with your plot
```

Write 2-3 sentences describing your observations and what this data might suggest (or not suggest) about the impact of smoking on birth weight.

> *REPLACE THIS TEXT WITH YOUR ANSWER*

**(d)** Construct a plot that shows the relationship between birth weight (`weight`) and gestational age (`weeks`). Make sure to specify meaningful axis labels where appropriate.

```{r, message=FALSE, fig.width=3, fig.height=4, fig.align='center'}
# To figure out how to do this google "ggplot2 scatter plot", or check out
# - https://ggplot2.tidyverse.org/#usage
# - https://ggplot2.tidyverse.org/#cheatsheet 
# - https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf
```

Write 2-3 sentences describing your observations and what this data might suggest (or not suggest) about the impact of gestational age on birth weight. Does this change the interpretation of your results above?

> *REPLACE THIS TEXT WITH YOUR ANSWER*